Assessing United States county-level exposure for tropical cyclone impact research

### 2020-08-09 ehp_revision/response_to_reviewers.Rmd

Other approaches have also been suggested to reduce ecological bias while 
leveraging ecological data, which might be available at a much larger scale---and
so provide more power---than is practical in collecting individual data. These
include multi-level approaches in which individual-level data from a sample
of individuals in each community is incorporated.

The second issue is exposure measurement error/misclassification. When the 
exposure (e.g., binary exposure or continuous measure of intensity of exposure)
is only measured at the county level, all the individuals in that county are 
assigned that same exposure. However, the hazards included in this study will 
vary across each county. For some hazards, this variation will be more stark. 
For example, tornadoes can be very local, causing extreme damage to a row of 
homes and none to the homes across the street. Some types of floods can also 
be very localized, rather than creating widespread flooding throughout 
the county. In both cases, when a county-level exposure assessment is made
based on an event in the county in association with the storm, there may be
many individuals in the county who did not experience that hazard at their 
residence. The other hazards we included---storm-associated winds and rainfall---will
also have some variation across counties, but this will often not be as stark. 
The rainfields for tropical cyclones are very large, and while there are rainbands
within the storm that might have particularly high rates of precipitation, these
progess over the course of the storm, and it is unlikely that a county will have
one area that experienced very extreme precipitation while another experienced
very little. Winds might have more variation, especially in coastal counties and 
counties near the storm's central track, since winds decay quickly as you move
from the storm's center, and storms typically weaken rapidly once they make 
landfall and can no longer draw on warm ocean waters for energy.

This exposure data is provided at the county level, and this can lead to 
exposure measurement error / exposure misclassification in studies that 
infer to the individual level. In this case, there will be several processes
at play. First, the hazards of the storm will not be uniform across the
county. The degree of this within-county spatial heterogeneity will usually 
vary by hazard. The within-county heterogeneity is likely lowest for rain and for
wind at lower levels. For rain, there can be particularly severe rain bands, 
but often a storm's rainfield will bring high precipitation to a large 
area. Wind can have a high heterogeneity close to the storm's center 
(storm winds decay most quickly with distance near the storm's center)
and near landfall (winds decay rapidly after landfall). Other hazards can 
be very local within the county, only affecting a small proportion of the total
county area. A tornado is a clear example---while it causes severe destruction
in its track, areas just off the track can be unharmed, creating a dramatic
spatial gradient in damage within a county. Floods can follow both. Some hurricanes
have caused widespread flooding (e.g., Floyd, Katrina) while some have caused
flooding that is more spatially focused (inland, with flashfloods? coast flooding
just near the coast?). For tidewater/coastal areas, there is often a dense 
network of waterway. When rains occur further up the watershed, these all
become flooded.

Second is the question of whether exposure measured at the resident's location
would be ideal. This is the case typically when studying an exposure that is 
a chemical substance. In this case, the pathway depends on the substance entering 
the body. For tropical cyclones, health risk could result from stress, problems
accessing medical care, or repercussions of power outages (higher exposure to 
outdoor hazards like heat and air pollution, problems storing food and using 
some medical equipment). For these, high levels of wind and other hazards in the
person's community may be as or more important than at their precise location.

Measurement error can be either random or systematic. Systematic error can often 
be corrected with adjustment if the direction and typical size of the error is 
understood. Random error cannot in the same way. Either type of measurement error
can be either differential (associated with the probability of the outcome) or 
non-differential (independent of the distribution / probability of the outcome).
In simpler models, this characteristic might help in predicting whether the resulting
bias is likely toward the null; however, more complex models (e.g., statistical 
models with adjustment for potential confounders) are trickier to diagnose in terms
of the likely implications of differential versus non-differential measurement 
error [?].

Not only can the exposure be measured with error, but so too can potential confounders
and potential effect modifiers.

When a proxy exposure estimates is used for many people covered by a study, this
can result in a type of exposure measurement error called *Berkson error*. In this 
case, the true exposure of each individual is randomly distributed around this 
proxy/aggregate-level exposure estimate, with a mean exposure level among the 
subjects that is equal to the proxy/aggregate-level measure. In other words, the 
group as a whole is assigned a common exposure level, when in fact this is the 
average exposure level across members of that group, but the individuals' 
true exposure levels are randomly distributed around this mean group level to 
which they are all assigned for modeling.

Berkson:

$$
T = X + E
$$
where:

- $T$: true individual exposure
- $X$: proxy exposure, assigned to everyone in the group
- $E$: error, with mean of 0, independent of $X$

The dataset provides exposure by group/area (county), not by individual. These
ecological exposure estimates require us to assume that the exposure level is
the same for all individuals in that county. It ignores/smooths over within-county
variability in exposure.

For wind and rain, we're essentially basing exposure assessment on modeling of
exposure. For rain, this is a based on a re-analysis model that
incorporates/integrates observed data from the area and time, using a model to
integrate data from diverse sources. For wind, the exposure estimates are based
on a model, with a core based on the Willoughby wind model. For flooding and
tornadoes, the exposure is based on records of events anywhere in the county,
and in many cases may be based, essentially, on exposure assessment, as in many
cases someone like someone at a NOAA [?] center have determined whether the
event should be reported in the database.

Even exposure at each person's residence would not suffice, since people move 
around during the day (although perhaps less for most people during a severe
storm).

@@

There can also be an *individualistic fallacy*, if emphasis is limited to the
association between individual-level factors, without considering the potential
role of contextual factors in an individual's risk of a health outcome. In other
words, cross-level bias can go in both directions [@idrovo2011three]. While the
fallacy of inferring individual-level associations from ecologic-level data is
called *ecological fallacy*, the fallacy of inferring ecologic-level
(contextual) effects from individual-level data is known as the *atomistic
fallacy* [@idrovo2011three]. There are also fallacies that can come in, it
seems, from inferring at the same level as the data but failing to consider
influences from factors at the other level. When inferring individual-level
associations from individual-level data, without considering an additional role
of ecological-level factors, this is known as the *psychologistic* or
*individualistic fallacy* [@idrovo2011three]. When inferring contextual effects
from ecologic data without considering individual-level factors, this is known
as the *sociologistic fallacy* [@idrovo2011three].

> "Although ecological studies are important to epidemiology (especially in 
environmental and social epidemiology), public health practitioners seem 
afraid of ecological studies. It is a common practice to assume the presence of 
ecological fallacy (Robinson 1950) and low-level validity when analyzing an 
ecological study. Most epidemiologists prefer an exclusive individualistic
approach, although the importance of a multilevel causal approach is widely 
recognized (Diez-Roux 2002). In this sense, some authors suggest that it is
as important to recognize the presence of ecological fallacy as to 
recognize psychologistic or individualistic fallacy (Subramanian et al. 2009)."
[@idrovo2011three]

Confounding in general:

> "A confounder ... is associated with an exposure or risk factor for the outcome
and with the outcome independent of the exposure, but is not on the causal pathway
between the exposure and the outcome. Confounding occurs when the relationship
between an exposure variable and the outcome variable is contaminated, so that the
measure of association between these two variables is actually also capturing the
effects of a third variable, the confounding variable. For example, many factors 
are associated with income (exposure) and health (outcome), such as education, 
employment or wealth? It may be that these factors explain part of the observed
assocation." [@gunasekara2008glossary]

Measurement error in general:

> "Measurement error: The different between an observed variable and the variable 
that belongs in a multiple regression equation. Measurement error is the random 
or systematic error arising during data collection of variables. The measured
variable x is measured with error eta, which is the distance from x to the 
'true' value of x. Measurement error can produce bias such as attenuating the
estimators of exposure variables, but may also have more complex effects. Longitudinal
data analyses such as fixed effects models can actually augment the problem of 
measurement error in mis-measured explanatory variables that change little over
time." [@gunasekara2008glossary]

Study validity:

> "Study validity refers to the degree to which the inferences drawn from a study are
warranted when account is taken of the study methods; the representativeness of
the study sample; and the nature of the population from which it is drawn.[5]
There are two types of study validity. *Internal validity* is the degree to which the
results of a study are correct for the sample of people being studied. *External 
validity* (generalisability) is the degree to which the study results hold true
for a population beyond the subjects in the study or other settings.[20]"

@@

Therefore, it would be relatively straightforward to expand the wind exposure data
to cover other countries affected by Atlantic-basin storms, although this would 
require some expansion of the area covered by our current landmask. More modifications
would need to be made to the wind model software before it could be extended
to storms in other basins, particularly basins that use different wind averaging 
periods for their main tracking data and basins in the Southern Hemisphere. 
Alternatively, other wind modeling software could be used for this exposure 
assessment, as other groups have developed software that is more extensible to 
multiple storm basins [Australian group's software].

The rain exposure data would also be fairly easy to extend to cover other countries, 
although again there would be a few practical limitations to extend the approach
we've taken in developing the dataset covered in this manuscript. For this exposure, 
we've used a re-analysis dataset. We've taken some steps to translate this gridded data
to the county scale, so it can be integrated with the scale for many human impacts
measurements. However, these steps have been well documents [ref] and would be easy to
reproduce. The re-analysis dataset that we use is specific to the contiguous US [ref
on NLDAS], but other similar re-analysis datasets exist for other locations [example
of one with global coverage?].

For all of these, we have provided exposure assessments at the county scale. Any of the 
steps we have taken could be done for other countries at a similar geopolitical scale, 
although the name of that areal unit will vary by country (e.g., municipalities in Mexico, 
districts in India). For wind and rain exposures, the approach we take here could be
used with any areal unit, as long as there are geographical shape files available for
the area boundaries (although the modeling for wind may be to the area's geographic
mean center rather than population mean center).

> "We initially developed software to model exclusively Atlantic-basin storms
[@stormwindmodel]. There are two characteristics of Northwestern Pacific Ocean
basin storm data that prevent direct use of the current version of the software.
First, the software currently requires Western Hemisphere longitude inputs.
Second, the ``Best Tracks'' data from the China Meteorological Administration
include central tropical cyclone intensity using a 2-minute averaging period,
while the software currently assumes that central intensities are given for a
1-minute averaging period, as is the case for storm track data from the US
National Hurricane Center. With the assistance of Ms. Schumacher, we will update
our tropical cyclone wind model software to appropriately handle 'Best Tracks' data from
the China Meteorological Administration, incorporating conversion factors
[@harper2010guidelines] to convert central storm maximum sustained wind
measurements to a common averaging period prior to applying further steps of the
wind model."

@@

> "Because of the historic lack of direct measurements and remotely sensed tornado 
wind speeds at or near group level, damage surveying has remained the most common 
form for indicating tornado strength. ... The occurrence of a direct tornado strict 
upon a fixed, sufficiently sturdy, and well-calibrated wind measuring station is quite
rare. Only 31 direct in situ tornado observations are evident between 1894 and 2011."
[@edwards2013tornado]

\begin{shaded}
\textbf{R1 Comment 4:}
Line 95---wondering if there is a better word for "hits" here. Maybe "exposed?"

\textit{Lines 92--95 in the original manuscript:}

\begin{quotation}
\noindent
\textit{"In many cases, these studies analyze multi-year, multi-community data,
allowing them to estimate average associations over many disasters and to
explore how a disaster's characteristics, or the characteristics of the
community it hits, modify associated health risks (e.g., Anderson and Bell 2010;
Son et al. 2012; Liu et al. 2017)."}
\end{quotation}
\end{shaded}

@@

**Response:**

Thank you for this suggestion. We have changed the wording in this sentence:

> **Text in revised manuscript (relevant change in bold):**
"In many cases, these studies analyze multi-year, multi-community data, allowing
them to estimate average associations over many disasters and to explore how a
disaster's characteristics, or the characteristics ofthe **affected
communities**, modify associated health risks (e.g., Anderson and Bell 2010; Son
et al. 2012; Liu et al. 2017)."

@@

We have moved other suggestions that you sent in "Track Changes" of the Word 
document for the original manuscript and address each in detail below.

**Response:**

\begin{shaded}
\textbf{Reviewer 2 (R2) Comment 1:}
Line 164: Isn't it obvious to use better exposure measures than proxies? Wind,
rainfall, tornado, and flood are direct products of storms while distance is
only proxy of those exposures. Rather than calculating agreement among the
exposures, it would be important to note that these are independent hazards that
can come together or come with different intensity and at different time
windows.

@@

> "Under the extremely unlikely scenario of extreme collinearity [(abs. rho of 0.999 
or higher)], the numerical matrix inversion required for estimation fails, resulting
in unstable and biased predictors." [@schisterman2017collinearity]

> "Only for extreme values of rho cloes to 1 or -1 is beta-squareed nonidentifiable
due to numerical instability; otherwise, the correctly specified model produces
the most desirable (e.g., unbiased) estimates of the total effect of the exposure
on the outcome." [@schisterman2017collinearity]

Collinearity can cause imprecision in coefficient estimates, even if those
estimates are unbiased [@schisterman2017collinearity].

> "Multicollinearity: A term that refers to correlation among the independent
variables in a multiple regression model; it is usually invoked when some 
correlations are 'large', but an actual magnitude test is not well defined.[1]"
[@gunasekara2008glossary]

> "Collinearity occurs in multiple regression models when two (or more---multicollinearity)
exposure variables are included in the model but are so similar to each other that
they are essentially measuring at least part of the same thing. Collinearity 
creates problems with interpreting the analysis by affecting the standard errors
of the variables (as it increases variance)[24] and by causing biased estimates for one
or both of the collinear terms (sometimes dramatically so) [25]. If there is perfect
collinearity between two variables then one should be dropped from the model. However, 
dropping exposure variables to reduce multicollinearity may lead to bias in the model
from dropping useful information." [@gunasekara2008glossary]

> "It might be argued in some cases that it is appropriate to include (collinear) 
variables in a model even when they are highly correlated, because the theoretical 
basis for including the variables is strong and the results of the model are consistent
with expectations [10]. In such cases, models with and without the variables would be
tested and compared. Such consideraion of alternative models tends to be described
as an issue of (mis)-specification in econometrics. However, in epidemiology this
tends to be described in terms of mediating and *intermediary variables*. Thus, an
epidemiologist might argue for including an intermediate variable in a model (and 
testing it against a model without the variable) because including that variable might 
elucidate direct and indirect pathways between the exposure and outcome variables. 
An econometrician might do the same thing but describe this in terms of finding the
correct model specification." [@gunasekara2008glossary]

@@

> **Text in original manuscript:**
"The rainfall data are generally well-correlated with ground-based observations,
but may sometimes underestimate very high rainfall values (Figure 2). When
rainfall data is used to create binary exposure classifications, this
disagreement is unlikely to influence results, as both data sources agree in
identifying these as storms with high rainfall, but would be important to
consider for cases that include rainfall as a continuous measurement."

We have added estimates of these correlations specifically for high-precipitation
events in a new table in the Supplemental Material (Table S1, reproduced 
as Table \@ref(tab:highprecipcorr) of this response document).

\input{tables/precip_high_corr.tex}

\begin{shaded}
\textbf{R2 Comment 4:}
Figure 4: I would add Y-axis label "NOAA classification."
\end{shaded}

@@

\textit{Line 491--495 from original manuscript:}

\begin{quotation}
\noindent
\textit{"For another set of tropical cyclones (e.g., Ernesto in 2006, and Bertha
in 1996, Isabel in 2003), there was moderate to good agreement for pairwise
combinations of distance, rain, and wind, but poor agreement for other
combinations of metrics, while for another set of storms (e.g., Matthew in 2004
and Katrina in 2005), there was moderate to good agreement between distance and
rain."}
\end{quotation}
\end{shaded}

**Response:**

> "A job exposure matrix (JEM) may be defined as a cross classification of jobs
and occupational exposures. Some matrices are based on tasks, instead of jobs. 
... Usually there is a group of experts who build the matrix based on their 
own knowledge and experience on occupational exopsures. The experts should 
assess exposure to an open or fixed list of occupational exposures (that can 
include chemical, physical, biological, and/or psychosocial agents) for an
open or fixed list of jobs (or tasks) usually coded according to some established
national or international classification system. Some JEMs also include 
information for different calendar time periods." [@garcia2003glossary]

### 2020-08-09 ehp_revision/response_to_reviewers.Rmd

Second is the question of whether exposure measured at the resident's location
would be ideal. This is the case typically when studying an exposure that is a
chemical substance. In this case, the pathway depends on the substance entering
the body. For tropical cyclones, health risk could result from stress, problems
accessing medical care, or repercussions of power outages (higher exposure to
outdoor hazards like heat and air pollution, problems storing food and using
some medical equipment). For these, high levels of wind and other hazards in the
person's community may be as or more important than at their precise location.

Measurement error can be either random or systematic. Systematic error can often
be corrected with adjustment if the direction and typical size of the error is
understood. Random error cannot in the same way. Either type of measurement
error can be either differential (associated with the probability of the
outcome) or non-differential (independent of the distribution / probability of
the outcome). In simpler models, this characteristic might help in predicting
whether the resulting bias is likely toward the null; however, more complex
models (e.g., statistical models with adjustment for potential confounders) are
trickier to diagnose in terms of the likely implications of differential versus
non-differential measurement error [?].

Not only can the exposure be measured with error, but so too can potential
confounders and potential effect modifiers.

We provide exposure assessments at the county level. This spatial scale allows
for easy integration with health outcome data aggregated at the county level.
However, there are also some things to consider when modeling associations
at an ecological rather than individual level. The considerations are 
particularly nuanced when investigating the health risks associated with a 
disaster.

Often in epidemiology, a study aims to estimate how an exposure changes the
risk of a health outcome for an individual. The data might be collected for
lots of people, but the key question is this---if a person is exposed, how 
does his or her risk of the outcome change. If you are exposed to this hazard, 
how much more likely are you to have the outcome? To get sick, or injured, or
to die? This inference is at the individual level.

Even when individual-level inference is the goal, sometimes data is only 
available at an aggregated level. For example, privacy concerns might
mean that health outcomes are only made available as a county-wide daily 
count. You can fit a mode with this aggregated data, which will estimate the
association between exposure and health risk at the ecological level. This
contextual effect estimate is sometimes used as an estimate of individual-level
association, but doing so can introduce ecological/multi-level bias. The 
individual-level effect estimate can be very biased from the true association, 
even to the point of reversing the effect estimate---estimating a protective
effect when the true effect is detrimental, for example.

This ecological fallacy plays out in two ways. First, if you use a single 
exposure measurement for everyone in an area, some people will be misclassified
(or have exposure measured with error, for a continuous metric) unless the 
exposure is perfectly homogeneous across the area. This will typically not
be the case. This exposure misclassification can bias estimates of the 
association between exposure and outcome in the same way exposure 
misclassification through any other mechanism would.

Next, ecological bias could result from confounding, even if the confounders
are controlled in the ecological-level model. When you aggregate data, you 
lose information about how both the exposure and potential confounders vary
across individuals within the area of aggregation. As a result, a factor
could still confound the inference of an individual-level association, even 
if it is controlled for at a population level in an ecological model. For 
example, a study of the association between risk of pre-term birth and 
tropical cyclone exposure could control for county-level smoking when 
modeling county-level storm exposure and county-level rates of pre-term births.
Even with this control, an observed association could result from differences
in individual smoking status, if there is within-county variation in 
smoking and if this has a different pattern across people in the county than
variation in exposure from the county-wide exposure estimate.

For disasters, there are added nuances. First, for exposure it may not just 
matter what the level of exposure is in a person's immediate scope. Disasters
bring physical hazards that can harm people directly, but also through indirect
pathways. For example, a tropical cyclone can bring high winds that cause 
power outages, and as a result those affected could be exposed to more outdoor
hazards (outdoor air pollution, heat), struggle to safely store perishable
food and medications, and lose means to power medical equipment. While extreme
winds at a person's residence would increase their risk of a power outage, 
outages could also be caused by damage to the grid in another part of the 
community. In some cases, then, the level of exposure in a person's community
may be as important in opening a pathway of risk as exposure at the person's
immediate location.

What's more, if the disaster has a large health impact, the health outcome of
one person in the community could affect the risk of the outcome (or other
adverse outcomes) for others. This situation is often only the case for 
infectious diseases, where one person with the disease can spread it to others.
However, if the community-wide impact is large enough, it can affect access
to and effectiveness of medical care for everyone in the community. This
effect has been seen recently with Covid 19---attempts to "flatten the curve"
aim to avoid moving into a state where a community's health system becomes
overwhelmed and can no longer deliver a typical level of care to those in the
community. This effect could happen with either infectious or non-infectious
diseases.

@@

\begin{shaded}
\textbf{R2 Comment 8:}
Line 515: It would be informative to provide correlation coefficient for only
above 75 mm measures.

\textit{Line 514--515 from original manuscript:}

\begin{quotation}
\noindent
\textit{"The rainfall data are generally well-correlated with ground-based
observations, but may sometimes underestimate very high rainfall values (Figure
2)."}
\end{quotation}
\end{shaded}

**Response:** 
This is a great suggestion and helps provided quantitative support to the text
in the paper that the reviewer highlighted when making this comment. We have
added a table in the Supplemental Material that provides the Spearman
correlation coefficients for both all tropical cyclones to come near the sample
counties (replicating the analysis shown in Figure 2 of the main text) and
then when the analysis is limited to only those events with cumulative
precipitation of 75 mm or higher based on either data source.

In several cases, the correlation was much weaker when limiting analysis to
these severe rainfall events. For example, in Miami-Dade County, FL, the
Spearman correlation coefficient was 0.94 for all 64 tropical cyclones that came
near the county over the analysis period, while it was only 0.49 among the 18
storms with cumulative precipitation of 75 mm or more. There was a similar
reduction in Mobile County, AL. However, for some counties, the correlation
remained high (e.g., Harris County, TX; Orleans Parish, LA), although in all
cases there are small numbers of extreme precipitation events (5--20 across
these sample counties), and so there is likely substantial random variation in
these correlation estimates, since they are based on small numbers of 
observations.

[Include copy of table here.]

Finally, we have added a bit to the text that the reviewer cited to mention 
this point and to direct readers to the added supplemental table:

> Text in revised manuscript (relevant addition in bold):
"The rainfall data are generally well-correlated with ground-based
observations, but may sometimes underestimate very high rainfall values (Figure
2), **and in some counties the correlation was substantially lower when 
considering only tropical cyclones with cumulative local rainfall of 75 mm
or more (Supplemental Table 1)**."

@@

\begin{shaded}
\textbf{R2 Comment 9:}
Line 602: This is not a new finding, but how much misclassified would be a good
question to answer.

\textit{Line 600--602 from original manuscript:}

\begin{quotation}
\noindent
\textit{"Based on our results, the use of a distance-based metric to assess
exposure to any of these hazards, or the use of measurements from one hazard as
a proxy for exposure to any of the other hazards considered, would often
introduce exposure misclassification."}
\end{quotation}
\end{shaded}

**Response:** 
[Add new analysis of what percent of counties misclassified if distance 
is used as proxy for each hazard?]

@@

\begin{shaded}
\textbf{R2 Comment 10:}
Lines 626--627: It would still be beneficial to treat each exposure as a single
variable first and then consider overall impact. Only if each exposure data are
available.

\textit{Line 626--628 from original manuscript:}

\begin{quotation}
\noindent
\textit{"For these storms, it may be possible to assess exposure to multiple
hazards of the storm using a single metric, perhaps even a proxy like the
distance between the county and the storm’s track."}
\end{quotation}
\end{shaded}

**Response:** 
We agree. This type of approach has not always been easy in the past, but can
be easily implemented with the dataset we publish here. We have added to the
text on this point:

> Text in revised manuscript (relevant addition in bold):
"For these storms, it may be possible to assess exposure to multiple
hazards of the storm using a single metric, perhaps even a proxy like the
distance between the county and the storm’s track. **With the dataset we 
describe in this paper, however, there is little need to limit analysis
based on exposure to a single hazard or proxy, although multi-hazard studies
of storms with high agreement among hazard exposures should look out for 
modeling issues from multicollinearity.**"
